### 2013-09-07 exact_learning/paper.tex
spelling, slight refinement of exact learning...

These two observations can be used in a number of ways.
\citet{finley2008training} used 2. to show that learning with an
over-generating approach minimizes an upper bound on the empirical risk.

@@

\subsection{Combining Inference Procedures}\label{combining}
The cutting plane method described in Section~\ref{cutting_plane} relies only
on some separation oracle $I$ that produces violated constraints.

@@

In particular, we employ three strategies for producing constraints,
moving from one to the next if no more constraints can be found:
\begin{enumerate*}
    \item Produce a constraint using previous, cached inference results.
    \item Use a fast under-generating algorithm.
    \item Use exact inference.
\end{enumerate*}

While using more different oracles is certainly possible, we found
that using just these three methods performed very well in practice.  This
combination allows us to make fast progress initially and guarantee optimality
in the end.
Notably, it is not necessary for an algorithm used as 3) to always produce
exact results. For guaranteeing optimality of the model, it is sufficient that
we obtain a certificate of optimality when learning stops.

@@

For exact inference, we use the recently developed Alternating Direction Dual
Decomposition (AD$^3$) method of \citet{martins2011augmented}. AD$^3$ produces
a solution to the linear programming relaxation, which we use as the basis for
branch-and-bound.

@@

\subsection{Image Segmentation}
We use superpixel-based semantic image segmentation as our application as this is our task of interest. %TODO formulation
for this work.  CRF based models have a history of success in this application,
with much current work investigating models and
learning~\citep{gonfaus2010harmony, lucchi2013learning, ladicky2009associative,
kohli2009robust, krahenbuhl2012efficient}.  We use the MSRC-21 and Pascal VOC 20120 datasets, two
widely used benchmark in the field.

@@

\subsubsection{Pascal VOC 2010}
For the Pascal VOC 2010 dataset, we follow the procedure from \citet{krahenbuhl2012efficient}
in using the official ``validation'' set as our evaluation set, and splitting the training set again.
We use the unary potentials provided by the same work, and compare only against methods
using the same setup and potentials, \citet{krahenbuhlparameter} and \citet{dann2012pottics}.
Note that state-of-the-art approaches, some not build on the CRF framework, obtain
around 40\% Jaccard score (also call VOC score), notably \cite{xia2012segmentation}, who 
evaluate on the Pascal VOC 2010 ``test'' set.

@@

\subsubsection{Results}
We compare classification results using different inference schemes in
with results from the literature. As a sanity check, we
also provide results without pairwise interactions.

@@

The objective function values using only the under-generating move-making and
the exact inference are detailed in Table~\ref{msrc_objective} and Table~\ref{pascal_objective}.
We see that a significant gap between the cutting plane objective and the primal objective
remains when using only under-generating inference.
Additionally, the estimated primal objective $o^I$ using under-generating inference is
too optimistic, as can be expected. This underlines the fact that
under-generating approaches can not be used to upper-bound the primal
objective.

@@

\section{Summary}
In this chapter we demonstrated that it is possible to learn state-of-the-art CRF models
exactly using structural support vector machines, despite the model containing many loops.
The key to efficient learning is the combination of different inference mechanisms and
a novel caching scheme for the one-slack cutting plane method, in combination
with state-of-the-art inference methods.

We show that guaranteeing exact results is feasible in a practical setting, and
hope that this result provides a new perspective onto learning loopy models for
computer vision applications.

Even though exact learning does not necessarily lead to a large improvement in
accuracy, it frees the practitioner from worrying about optimization and
approximation issues, leaving more room for improving the model, instead of the
optimization.

While our model is trained on a relatively small dataset, we expect that scaling to
datasets of several thousand images, such as Pascal VOC, is unproblematic.
The time for a single iteration of the cutting plane solver will only grow
linearly for the inference. The cost of solving the QP is actually independent
of the number of samples, as only a single constraint will be added in each
iteration.

>>>>

@@

\subsection{Combining Inference Procedures}\label{combining}
The cutting plane method described in Section~\ref{cutting_plane} relies only
on some separation oracle $I$ that produces violated constraints when
performing loss-augmented prediction.

@@

In particular, we employ three strategies for producing constraints,
moving from one to the next if no more constraints can be found:
\begin{enumerate*}
    \item Produce a constraint using previous, cached inference results.
    \item Use a fast under-generating algorithm.
    \item Use a strong but slow algorithm that can certify optimality.
\end{enumerate*}

While using more different oracles is certainly possible, we found
that using just these three methods performed very well in practice.  This
combination allows us to make fast progress initially and guarantee optimality
in the end.
Notably, it is not necessary for an algorithm used as 3) to always produce
exact results. For guaranteeing optimality of the model, it is sufficient that
we obtain a certificate of optimality when learning stops.
% TODO certify optimality != exact...

@@

For inference with optimality certificate, we use the recently developed
Alternating Direction Dual Decomposition (AD$^3$) method of
\citet{martins2011augmented}. AD$^3$ produces a solution to the linear
programming relaxation, which we use as the basis for branch-and-bound.

@@

\subsection{Semantic Image Segmentation}
% TODO reference auf davor / intro?
Our main application is of course semantic segmentation and object class
segmentation. We evaluate the proposed learning approach on Pascal VOC 2013 and
MSRC-21

@@

\subsubsection{Pascal VOC 2010}
For the Pascal VOC 2010 dataset, we follow the procedure from \citet{krahenbuhl2012efficient}
in using the official ``validation'' set as our evaluation set, and splitting the training set again.
We use the unary potentials provided by the same work, and compare only against methods
using the same setup and potentials, \citet{krahenbuhlparameter} and \citet{dann2012pottics}.
Note that state-of-the-art approaches, some not build on the CRF framework, obtain
around a Jaccard Index (also call VOC score) of 40\% , notably \cite{xia2012segmentation}, who 
evaluate on the Pascal VOC 2010 ``test'' set.

@@

\subsubsection{Results}
We compare classification results using different inference schemes with
results from the literature. As a sanity check, we also provide results without
pairwise interactions.

@@

The objective function values using only the under-generating move-making and
the exact inference are detailed in Table~\ref{msrc_objective} and Table~\ref{pascal_objective}.
We see that a significant gap between the cutting plane objective and the primal objective
remains when using only under-generating inference.
Additionally, the estimated primal objective $o^I$ using under-generating inference is
too optimistic, as can be expected. This underlines the fact that
under-generating approaches can not be used to upper-bound the primal
objective or compute meaningful duality gaps.

@@

\section{Summary}
In this chapter we demonstrated that it is possible to learn state-of-the-art
conditional random field models exactly using structural support vector
machines, despite the model containing many loops.  The key to efficient
learning is the combination of different inference mechanisms and a novel
caching scheme for the one-slack cutting plane method, in combination with
state-of-the-art inference methods.

We show that guaranteeing exact results is feasible in a practical setting, and
hope that this result provides a new perspective onto learning loopy models for
computer vision applications.
%
Even though exact learning does not necessarily lead to a large improvement in
accuracy, it frees the practitioner from worrying about optimization and
approximation issues, leaving more room for improving the model, instead of the
optimization.

### 2013-09-07 structured_prediction/paper.tex
spelling, slight refinement of exact learning...

This approach to structured prediction is closely related to 
approachs using \emph{probabilistic graphical models}.  Probabilistic graphical models are a
tool to express factorizations of probability distributions.  Similar to
\Eqref{log_factor_graph}, the joint probability distribution over a
multi-variate random variable $y$ can be expressed using a factor-graph:
\begin{equation}\eqlabel{graphicalmodel}
    p(y | x) = \frac{1}{Z_x}\prod_{F \in \mathcal{F}} \exp\left(\psi_F(x, y_{N_F})\right)
\end{equation}
where
\begin{equation}
    Z_x = \sum_{y' \in \mathcal{Y}} \prod_{F \in \mathcal{F}} \exp\left(\psi(x, y'_{N_F})\right)
\end{equation}
is the normalization constant of the conditional distribution over $y$.
If $f$ is chosen as in \Eqref{linear_psi}, then the resulting distribution
belongs to the exponential family.

@@

Additionally, we discuss practical implications and implementation.  One
particularly intereresting aspect is the difference between \emph{sequention}
(or \emph{online}) and \emph{batch} algorithms.  Batch algorithms process the
whole dataset before adjusting parameters, while sequential algorithms process
one sample at a time, and adjust parameters incrementally.
In image segmentation tasks, inference is often costly, making loss-augmented
prediction the most expensive step in learning. This often leads to longer
run-times for batch algorithms. On the other hand, loss-augmented prediction in
batch algorithms is embarassingly parallel, allowing the use of multiple
processors with almost linear speed improvements.

@@

Arguably the most straight-foward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
to work in a stochastic setting (see \citet{ratliff2007online}).
Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
\end{align}
The most commonly used update has the simple form
\begin{equation}
    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of step sizes.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
\citet{shalev2011pegasos} showed that this schedule achieves a convergence rate of $O(\frac{\ln T }{T})$.
\citet{lacoste2012block} and \citet{shamir2012stochastic} recently showed independently that
a rate of $O(\frac{1}{T})$ can be achieved using a novel averaging scheme:
\begin{equation}
    \bar{\theta}_{T} = \frac{2}{(T+1)(T+2)} \sum_{t=0}^T(t+1) \theta_t.
\end{equation}
This $t$-weighted averaging can be computed on-the-fly as
\begin{equation}\eqlabel{averaging}
    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}
\end{equation}

@@

The $n$-slack cutting plane is a sequential algorithm that processes each sample
individually. While this allows fast process of the optimization with respect
to the number of calls to loss-augmented prediction, individual steps
become more and more costly. The number of the constraints is usually a
multiple of the number of training  samples, which leads to very large QP
problems, even for medium sized datasets. This makes the algorithm often slow
in practice.
%
Solving the quadratic programm can be accelerated using several techniques. We
found that agressively removing constraints that are inactive or contribute
little to the solution often makes the difference between the algorithm being
practical or not.  Another possible technique is to update the quadratic
program only every $r$ samples, for some small integer $r$
(\citet{joachims2009cutting} suggest 100). While this strategy on its own did
not provide a large benefits in our experiments, it allows for parallel
loss-augmented prediction on these mini-batches of size $r$.

@@

\citet{joachims2009cutting}, who introduced the method, proposed two enhancements
to make the algorithm more efficient:
\begin{description}
    \item[Constraint Pruning] As in the $n$-slack algorithm, members of the
        working set $\W$ can become inactive during learning.  If a constraint
        has been inactive for a number of iterations, it is removed from $\W$,
        leading to smaller problem sizes.
    \item[Inference Caching] In the $1$-slack algorithm, each constraint is
        created using a combination of loss-augmented prediction results.
        Therefore, each of these predictions can be part of multiple
        constraints during learning.
        To exploit this, we maintain a set $C^i$ of the last $r$ results of
        loss-augmented prediction for each training example $(x^i, y^i)$
        (line~\ref{get_cutting_plane} in Algorithm~\ref{alg_one_slack}).
        For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we
        find
        \[ \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n
            \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\Phi(x^i, y^i) -
                \Phi(x^i, \hat{y})] \]
        by enumeration of $C^i$ and continue until
        line~\ref{convergence_check}.  Only if $\xi' - \xi < \epsilon$, that is
        the produced constraint is not violated stronly enough, we return to
        line~\ref{get_cutting_plane} and actually invoke the loss augmented
        prediction $I$.
\end{description}

>>>>

This approach to structured prediction is closely related to 
approaches using \emph{probabilistic graphical models}.  Probabilistic graphical models are a
tool to express factorizations of probability distributions.  Similar to
\Eqref{log_factor_graph}, the joint probability distribution over a
multi-variate random variable $y$ can be expressed using a factor-graph:
\begin{equation}\eqlabel{graphicalmodel}
    p(y | x) = \frac{1}{Z_x}\prod_{F \in \mathcal{F}} \exp\left(\psi_F(x, y_{N_F})\right)
\end{equation}
where
\begin{equation}
    Z_x = \sum_{y' \in \mathcal{Y}} \prod_{F \in \mathcal{F}} \exp\left(\psi(x, y'_{N_F})\right)
\end{equation}
is the normalization constant of the conditional distribution over $y$.
If $f$ is chosen as in \Eqref{linear_psi}, then the resulting distribution
belongs to the exponential family.

@@

Additionally, we discuss practical implications and implementation.  One
particularly interesting aspect is the difference between \emph{sequential}
(or \emph{online}) and \emph{batch} algorithms.  Batch algorithms process the
whole dataset before adjusting parameters, while sequential algorithms process
one sample at a time, and adjust parameters incrementally.
In image segmentation tasks, inference is often costly, making loss-augmented
prediction the most expensive step in learning. This often leads to longer
run-times for batch algorithms. On the other hand, loss-augmented prediction in
batch algorithms is embarrassingly parallel, allowing the use of multiple
processors with almost linear speed improvements.

@@

Arguably the most straight-forward way to approach \Eqref{learning_equation} is
using subgradient descent.  In light of the complexity of solving the
loss-augmented prediction problem in \Eqref{loss_augmentation}, is is natural
to work in a stochastic setting (see \citet{ratliff2007online}).
Given a model through $\phi$ and a set of parameters $\theta$, a subgradient
considering a single training example $(x\hoch{i}, y\hoch{i})$
can be computed simply by solving the loss-augmented prediction problem:
\begin{align}
    &\frac{d}{d\theta} \left [ \frac{1}{2} ||\theta||^2 + C \ell(x\hoch{i}, y\hoch{i}, \theta) \right ] \ni C\left [ \phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i}) \right]+ \theta\\
    &\text{with }\hat{y} \in \argmax_{y \in \mathcal{Y}} \Delta(y\hoch{i}, y) + \theta^T \phi(x\hoch{i}, y)
\end{align}
The most commonly used update has the simple form
\begin{equation}
    \theta_{t+1} = (1 - \eta_t) \theta_t - \eta_t C \left [\phi(x\hoch{i}, \hat{y}) - \phi(x\hoch{i}, y\hoch{i})\right]
\end{equation}
Where $\eta_t$ is a sequence of step sizes.
%Theoretical convergence guarantees of \ldots can be given whenever \ldots,
but in practice the choice of $\eta_t$ is often strongly influences the convergence behavior.
Many practitioners adopted the sequence proposed for binary SVMs in the Pegasos algorithm~\citep{shalev2011pegasos}:
\begin{equation}
    \eta_t = \frac{C}{t},
\end{equation}
which has been found to work well in many settings.
\citet{shalev2011pegasos} showed that this schedule achieves a convergence rate of $O(\frac{\ln T }{T})$.
\citet{lacoste2012block} and \citet{shamir2012stochastic} recently showed independently that
a rate of $O(\frac{1}{T})$ can be achieved using a novel averaging scheme:
\begin{equation}
    \bar{\theta}_{T} = \frac{2}{(T+1)(T+2)} \sum_{t=0}^T(t+1) \theta_t.
\end{equation}
This $t$-weighted averaging can be computed on-the-fly as
\begin{equation}\eqlabel{averaging}
    \bar{\theta}_{t} = \frac{t}{t + 2} \bar{\theta}_t + \frac{2}{t+2} \theta_{t+1}
\end{equation}

@@

The $n$-slack cutting plane is a sequential algorithm that processes each sample
individually. While this allows fast process of the optimization with respect
to the number of calls to loss-augmented prediction, individual steps
become more and more costly. The number of the constraints is usually a
multiple of the number of training  samples, which leads to very large QP
problems, even for medium sized datasets. This makes the algorithm often slow
in practice.
%
Solving the quadratic program can be accelerated using several techniques. We
found that aggressively removing constraints that are inactive or contribute
little to the solution often makes the difference between the algorithm being
practical or not.  Another possible technique is to update the quadratic
program only every $r$ samples, for some small integer $r$
(\citet{joachims2009cutting} suggest 100). While this strategy on its own did
not provide a large benefits in our experiments, it allows for parallel
loss-augmented prediction on these mini-batches of size $r$.

@@

\citet{joachims2009cutting}, who introduced the method, proposed two enhancements
to make the algorithm more efficient:
\begin{description}
    \item[Constraint Pruning] As in the $n$-slack algorithm, members of the
        working set $\W$ can become inactive during learning.  If a constraint
        has been inactive for a number of iterations, it is removed from $\W$,
        leading to smaller problem sizes.
    \item[Inference Caching] In the $1$-slack algorithm, each constraint is
        created using a combination of loss-augmented prediction results.
        Therefore, each of these predictions can be part of multiple
        constraints during learning.
        To exploit this, we maintain a set $C^i$ of the last $r$ results of
        loss-augmented prediction for each training example $(x^i, y^i)$
        (line~\ref{get_cutting_plane} in Algorithm~\ref{alg_one_slack}).
        For generating a new constraint $(\hat{y}^1, \dotsc, \hat{y}^n)$ we
        find
        \[ \hat{y}^i \leftarrow \argmax_{\hat{y}\in C^i} \sum_{i=1}^n
            \Delta(y^i, \hat{y}) - \theta^T \sum_{i=1}^n [\Phi(x^i, y^i) -
                \Phi(x^i, \hat{y})] \]
        by enumeration of $C^i$ and continue until
        line~\ref{convergence_check}.  Only if $\xi' - \xi < \epsilon$, that is
        the produced constraint is not violated strongly enough, we return to
        line~\ref{get_cutting_plane} and actually invoke the loss augmented
        prediction $I$.
\end{description}
